A Survey on Ordered Binary Decision Diagrams

### 2015-09-11 paper.tex
Add function argument omittance notice
Finish Restrict operation; add time complexity to Apply operation

This paper serves as survey on symbolic Boolean function manipulation using OBDDs.
The following section describes the representation of Boolean functions through
BDDs as well as OBDDs. It also highlights the advantages and disadvantages of the
OBDD representation. The third section describes operations which can be applied
to OBDDs efficiently. No source code is provided but those can be looked up in
the corresponding papers easily (references are provided for each algorithm).
Section~\ref{sec:implementation-aspects} describes aspects one has to take into
account when implementing an OBDD package, such as what data structures to use
or how to deal with the problem of variable ordering. Limitations of the OBDD
representation are described in the subsequent section. The last section
discusses alternative representations to possibly overcome the limitations of
the OBDD representation. \tbr

@@

\textit{Binary decision diagrams} (\textit{BDDs}) have been well studied since
Lee~\cite{LEE59} and especially Akers~\cite{AKERS78} introduced them. BDDs
represent a Boolean function $f$ as an acyclic, directed graph (DAG). Nodes
represent variables and have one incoming and two outgoing edges, each of which
denotes the variable evaluates to 0 (left) or 1 (right), respectively. Whenever
the result of $f$ is fixed, an edge points to one of the constant functions
\textbf{0} or \textbf{1}. To evaluate $f$ for a given variable assignment, one
just follows the path for the respective variable values. As soon as the path
ends up in a constant, i.e. \textbf{0} or \textbf{1}, this is the result of the
corresponding variable assignment. Akers~\cite{AKERS78} already provided some
techniques to simplify the BDD representation such as combining equivalent nodes.

@@

The following paragraphs summarize those operations and their properties briefly.
Implementation details are covered in Section~\ref{sec:implementation-aspects}.

@@

@@

By applying the three transformation rules of the reduction procedure in each
step, \textit{Apply} generates a reduced graph immediately~\cite{BRYANT92}.

There are some refinements when it comes to implementation. Those are covered in
Section~\ref{sec:implementation-aspects} as well as the original paper of
Bryant~\cite{BRYANT86, BRYANT92}. \tbr

@@

\section{Implementational Aspects}
\label{sec:implementation-aspects}

>>>>

This paper serves as survey on symbolic Boolean function manipulation using OBDDs.
The following section describes the representation of Boolean functions through
BDDs as well as OBDDs. It also highlights the advantages and disadvantages of the
OBDD representation. The third section describes operations which can be applied
to OBDDs efficiently. No source code is provided but those can be looked up in
the corresponding papers easily (references are provided for each algorithm).
Section~\ref{sec:implementational-aspects} describes aspects one has to take into
account when implementing an OBDD package, such as what data structures to use
or how to deal with the problem of variable ordering. Limitations of the OBDD
representation are described in the subsequent section. The last section
discusses alternative representations to possibly overcome the limitations of
the OBDD representation. \tbr

@@

\textit{Binary decision diagrams} (\textit{BDDs}) have been well studied since
Lee~\cite{LEE59} and especially Akers~\cite{AKERS78} introduced them. BDDs
represent a Boolean function $f$ as an acyclic, directed graph (DAG). The function
arguments may be omitted in some cases to keep the formulas simple, i.e. $f$
denotes $f(x_1, \ldots, x_2)$. Nodes represent variables and have one incoming
and two outgoing edges, each of which denotes the variable evaluates to 0 (left)
or 1 (right), respectively. Whenever the result of $f$ is fixed, an edge points
to one of the constant functions \textbf{0} or \textbf{1}. To evaluate $f$ for a
given variable assignment, one just follows the path for the respective variable
values. As soon as the path ends up in a constant, i.e. \textbf{0} or \textbf{1},
this is the result of the corresponding variable assignment. Akers~\cite{AKERS78}
already provided some techniques to simplify the BDD representation such as
combining equivalent nodes.

@@

The following paragraphs summarize those operations and their properties briefly.
Implementation details are covered in Section~\ref{sec:implementational-aspects}.

@@

Although \textit{Apply} is the core operation, the \textit{Restrict} operation is
discussed first, mainly because it is used within \textit{Apply}. Besides,
\textit{Restrict} is straightforward and easy to understand.

A variable $x$, argument of a function $f$, can be restricted to a given value
$c$, denoted by $f|_{x\leftarrow c}$, as follows.

Starting from the root $n$ of the OBDD representation of $f$ and traversing the
OBDD depth-first, every node is checked for $var(n) = x$. This is done recursively
on the given graph, generating the resultant graph This node represents
the variable to be restricted. Now, there are two cases for $c$ to distinguish:
\begin{enumerate}
    \item{$c=0$: The incoming edge of $n$ is redirected to $lo(n)$.}
    \item{$c=1$: The incoming edge of $n$ is redirected to $hi(n)$.}
\end{enumerate}

By applying the three transformation rules of the reduction procedure in each
step, \textit{Restrict} generates a reduced graph immediately~\cite{BRYANT92}.

This procedure has linear time complexity in the number of nodes, i.e. $O(|V|)$,
assuming an efficient implementation (see
Section~\ref{sec:implementational-aspects} for details)~\cite{BRYANT92}. 
\tbr

@@

To get a reduced graph immediately, the three transformation rules are utilized
again for each step~\cite{BRYANT92}.

There are some refinements when it comes to implementation. Those are covered in
Section~\ref{sec:implementational-aspects} as well as the original paper of
Bryant~\cite{BRYANT86, BRYANT92}.  However, assuming an efficient implementation,
the time complexity can be bound to
$O\left(|V_{f_1}| \cdot |V_{f_2}|\right)$~\cite{BRYANT86, BRYANT92}. \tbr

@@

\section{Implementational Aspects}
\label{sec:implementational-aspects}
